---
title: "PAIC simulation results"
output: pdf_document
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r, warning=FALSE, include=FALSE}
library(data.table)
library(ggplot2)
library(dplyr)
library(tidyr)
library(stringr)
library(ggthemr)
library(patchwork)
library(data.table)
ggthemr::ggthemr("pale")
```


# Overview of the problem

The goal is to estimate the relative effect of treatment $A$ vs $B$, noted $\Delta_{t}$. These two treatments have been evaluated in two distinct trials, $AB$ and $AC$. The target estimand is the relative effect of $A$ vs $B$ in the source population of trial $AC$, written $\Delta_{t(AC)}$. 

$\Delta_{t (AB)}$ will be estimated using different estimators: 

- Unadjusted: 
  - Unanchored: naive model
  - Anchored: difference of relative effects $A$ versus $C$ and $B$ versus $C$
- Adjusted
  - Adjusted using a multivariable outcome model (regression model), or univariable weighted outcome model (propensity score)
    - Each of these models will be fitted using either Individual Patient Data (IPD) for both trials, or IPD for the trial $AB$ and AgD for trial $BC$: Population-Adjusted Indirect Comparisons (PAIC). PAIC multivariable outcome model = Simulated Treatment Comparison (STC), univariable weighted outcome model = Matching-Adjusted Indirect Comparisons (MAIC).
    - Each of these models will be either anchored or unanchored.

## Summary of the estimators

|Estimand|Common comparator?|Full IPD?|Estimator|
|--------|------------------|---------|---------|
|Marginal|Yes|Yes|NMA with IPTW PS|
|Marginal|No|Yes|IPTW PS|
|Marginal|Yes|No|Anchored MAIC|
|Marginal|No|No|Unanchored MAIC|
|Conditional|Yes|Yes|IPD NMA|
|Conditional|No|Yes|Regression-based outcome model|
|Conditional|Yes|No|Anchored STC|
|Conditional|No|No|Unanchored STC|


## (Current) scenario settings 

- Sample size per treatment arm = 100

- One gaussian outcome $Y$

- Treatment effects: 
 
  - $\beta_A = 1.5$  
  
  - $\beta_B = 1$  
  
  - $\beta_C = 0$(*no effect, ~ placebo*)

- One binary confounding factor $X_1$, following a binomial distribution $B(X_1)\sim 0.5$ in an overarching population from which patients of trials $AB$ and $AC$ have been sampled. Effects of this confounding factor on 1. trial assignment and 2. the outcome $Y$:

  1. Constant effect on the outcome (prognostic): $\beta_{Y_{X1}} = 1.5$, and effect on the outcome through interaction with treatment $A$ (interaction): $\beta_{Y_{A:X1}} = 1.2$
  
  2. Effect on trial assignment ($\beta_{AC, X1} = 2$) and ($\beta_{BC, X1} = 0$)
  
- Outcome model : $$Y \sim \left(\beta_{Y_{X_1:A}}A + \beta_{Y_{X1}}\right)X1 + \beta_{Y_A}A + \beta_{Y_B}B + \beta_{Y_C}C + \epsilon_Y$$

with $\epsilon_Y \sim N(0, 1)$

- Trial assignment model : $P(Z=AC) \sim g \left(\beta_{AC, X1} X_1 \right)$ with $g(X) = \left(1 + exp(-X)\right)^{-1}$

```{r, echo = TRUE, eval = FALSE}

prop_X1 <- 0.5
bT_X1 <- 2

bY_X1 <- 1.5
bYA_X1 <- 1.2

bY_A <- 1.5
bY_B <- 1.5
bY_C <- 0

N_RCT <- 200

trial_assignment_model_AC <- bquote(X1 * bT_X1)
trial_assignment_model_BC <- bquote(0) 
outcome_model <- bquote(bYA_X1*X1*A + bY_X1*X1 + bY_A*A + bY_B*B + bY_C*C)

```

## Monte-Carlo simulations settings

Two thousands iteration of that scenario, and estimation of variance of treatment effects via bootstrap (1000 iterations).

```{r Generating the overarching population, echo = TRUE, eval = FALSE}
library(data.table)
N_pop <- 10^6

pop_init <- data.table(
  id = 1:N_pop,
  X1 = rbinom(N_pop, 1, prop_X1)
)

trial_assignement_prob <- function(df, trial_assignment_model, deviates = FALSE) {
  predicted <- with(df, eval(trial_assignment_model))
  # variance of a logistic function is set ((s^2\pi^2)/3), with s=1 when using normal law -->
  # making an assumption about bias and residual heterogeneity
  return(plogis(predicted))
}

predict_outcome <- function(df, outcome_model, deviates = FALSE) {
  predicted <- with(df, eval(outcome_model))
  if (deviates) return(predicted + rnorm(n = length(predicted), mean = 0, sd = 1)) else return(predicted)
}

pop_init$prob_AC <- trial_assignement_prob(pop_init, trial_assignment_model_AC)
pop_init$prob_BC <- trial_assignement_prob(pop_init, trial_assignment_model_BC)


pop_init$Y_theo_A <- pop_init[, c("A", "B", "C") := .(1L, 0L, 0L)] |>
  predict_outcome(outcome_model)
pop_init$Y_theo_B <- pop_init[, c("A", "B", "C") := .(0L, 1L, 0L)] |>
  predict_outcome(outcome_model)
pop_init$Y_theo_C <- pop_init[, c("A", "B", "C") := .(0L, 0L, 1L)] |>
  predict_outcome(outcome_model)
pop_init[, c("A", "B", "C") := NULL]

pop_init$prob_AC <- trial_assignement_prob(pop_init, trial_assignment_model_AC)
pop_init$prob_BC <- trial_assignement_prob(pop_init, trial_assignment_model_BC)

ttt_names <- c("A", "B", "C")
cols_obs <- paste0("Y_obs_", ttt_names)
cols_theo <- paste0("Y_theo_", ttt_names)
# Assigning "observed" values for all the individual in the population, ie theoretical outcome + gaussian noise
pop_init[, eval(bquote(cols_obs)) := lapply(.SD, \(x) x + rnorm(n = length(x))), .SD = cols_theo]
```


```{r Sampling trials, echo = TRUE, eval = FALSE}
trial_AC <- pop_init[
  sample(pop_init$id, N_RCT, replace = FALSE, prob = pop_init$prob_AC), 
  .(id, ttt = factor(rep_len(c("A", "C"), length.out = N_RCT), levels = c("C", "A", "B")))][
    pop_init_long[,c("id", "X1", "ttt", "prob_AC", "theo", "obs")], on = .(id, ttt), nomatch = NULL
  ]
trial_BC <- pop_init[
  sample(pop_init$id, N_RCT, replace = FALSE, prob = pop_init$prob_BC),
  .(id, ttt = factor(rep_len(c("B", "C"), length.out = N_RCT), levels = c("C","A", "B")))][
    pop_init_long[,c("id", "X1", "ttt", "prob_BC", "theo", "obs")], on = .(id, ttt), nomatch = NULL
  ]

```


# Simulation results 

```{r}
results_simulations <- readRDS("results_simulations/results_simulations.RDS")
results_simulations_1 <- readRDS("C:/Users/aserret-larmande/git_repos/paic_simu/results_simulations/20231019_102445/experiment_1.RDS")
results_simulations_2 <- readRDS("C:/Users/aserret-larmande/git_repos/paic_simu/results_simulations/20231019_102445/experiment_2.RDS")
results_simulations_3 <- readRDS("C:/Users/aserret-larmande/git_repos/paic_simu/results_simulations/20231019_102445/experiment_3.RDS")
results_simulations_4 <- readRDS("C:/Users/aserret-larmande/git_repos/paic_simu/results_simulations/20231019_102445/experiment_4.RDS")
```


```{r}
long_results <- bind_rows(results_simulations, .id = "n_iter") |>
  mutate(data = ifelse(model %in% c("maic", "stc"),
                       "PAIC",
                       ifelse(
                         model %in% c("ml", "glm"),
                         "IPD",
                         ifelse(
                           model == "unadjusted", "AgD", NA)
                         )
                       ),
         data = factor(data, levels = c("AgD", "PAIC", "IPD")),
         anchored = factor(anchored, levels = c("unanchored", "anchored")),
         adjustment = ifelse(adjustment == "iptw", "IPTW", stringr::str_to_title(adjustment)),
         across(c(model, anchored), .fns = stringr::str_to_title)) |>
  rename_with(stringr::str_to_title)


true_conditional <- long_results |>
  filter(Adjustment == "True" & Model == "Conditional") |>
  distinct(Estimate) |>
  pull(Estimate)
true_marginal <- long_results |>
  filter(Adjustment == "True" & Model == "Marginal") |>
  distinct(Estimate) |>
  pull(Estimate)

df_true_effects <- data.frame(Adjustment = c("Unadjusted", "Regression", "IPTW"),
                              true = c(true_conditional, true_conditional, true_marginal))
```


## Distribution of the estimate 

```{r}
# Variation of the Estimate
long_results |>
  filter(Adjustment != "True") |>
  ggplot(aes(x = Estimate, y = Adjustment, color = Data, fill = Data)) +
  geom_violin(alpha = 0.8) +
  facet_wrap(~Anchored) +
  geom_vline(aes(xintercept = true_conditional, linetype = type),
             show.legend = TRUE,
             data = data.frame(effect = true_conditional, type = "dashed")) +
  scale_linetype_manual(values = c("dashed"), labels = c(str_wrap("True conditional effect", width = 10))) +
  labs(y = NULL,
       title = "Treatment effect estimates")
```

# Standard error of the estimate

```{r}
long_results |>
  filter(Adjustment != "True") |>
  ggplot(aes(x = Variance, y = Adjustment, color = Data, fill = Data)) +
  geom_violin(alpha = 0.8) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "black") +
  facet_wrap(~Anchored) +
  labs(y = NULL,
       x = "Standard error of the estimate",
       title = str_to_title("Treatment effect estimate standard error"))

```


\newpage

# Performance of the estimators

```{r }
# Long indicators
get_bias <- function(obs, theo) {
  mean(obs - theo, na.rm = FALSE)
}
get_RMSE <- function(obs, theo) {
  sqrt(mean((obs - theo)**2, na.rm = FALSE))
}
get_VR <- function(obs, se_obs) {
  mean(se_obs, na.rm = FALSE) / sd(obs, na.rm = FALSE)
}
get_cov_95 <- function(coef, se, theo) {
  ub <- coef + qnorm(0.975)*se
  lb <- coef - qnorm(0.975)*se
  covered <- theo < ub & theo > lb
  mean(covered, na.rm = FALSE)
}

df_stats <- long_results |> filter(Adjustment != "True") |>
  left_join(df_true_effects, by = "Adjustment") |>
  group_by(Adjustment, Model, Anchored, Data) |>
  summarize(bias = get_bias(Estimate, true),
            rmse = get_RMSE(Estimate, true),
            vr = get_VR(Estimate, sqrt(Variance)),
            cov_95 = get_cov_95(Estimate, sqrt(Variance), true), 
            .groups = "drop") |>
  pivot_longer(cols = c("bias", "rmse", "vr", "cov_95"),
               names_to = "indicator", values_to = "values")
```


```{r}
df_stats |> 
  select(indicator, everything()) |> 
  mutate(Model = ifelse(Model == "Unadjusted", "Unadjusted", toupper(Model)),
         indicator = case_match(indicator, 
                                "bias" ~ "Bias", 
                                "rmse" ~ "RMSE", 
                                "vr" ~ "VR", 
                                "cov_95" ~ "Cov 95"), 
         values = format(round(values, 3), digits = 3, nsmall = 3, justify = "right")
  ) |> 
  arrange(indicator, desc(Adjustment), Data, desc(Anchored)) |> 
  knitr::kable()
```


```{r }
bias_plot <- ggplot(filter(df_stats, indicator == "bias"),
                    aes(x = values, color = Data, shape = Anchored)) +
  geom_point(aes(y = Adjustment), position = "jitter", size = 3) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  labs(title = "Bias")
rmse_plot <- ggplot(filter(df_stats, indicator == "rmse"),
                    aes(x = values, color = Data, shape = Anchored)) +
  geom_point(aes(y = Adjustment), position = "jitter", size = 3) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  labs(title = "RMSE")
vr_plot <- ggplot(filter(df_stats, indicator == "vr"),
                    aes(x = values, color = Data, shape = Anchored)) +
  geom_point(aes(y = Adjustment), position = "jitter", size = 3) +
  geom_vline(xintercept = 1, linetype = "dashed") +
  labs(title = "VR")
ci_plot <- ggplot(filter(df_stats, indicator == "cov_95"),
                    aes(x = values, color = Data, shape = Anchored)) +
  geom_point(aes(y = Adjustment), position = "jitter", size = 3) +
  geom_vline(xintercept = 0.95, linetype = "dashed") +
  labs(title = "95% coverage")

bias_plot + rmse_plot + vr_plot + ci_plot +
  plot_annotation(title = "Performance of the estimators") +
  plot_layout(guides = "collect") &
  theme(legend.position = 'bottom',
        legend.box = "vertical",
        axis.title = element_blank())
```

