---
title: "PAIC simulation results"
output: pdf_document
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r, warning=FALSE, include=FALSE}
library(ggplot2)
library(data.table)
library(dplyr)
library(tidyr)
library(stringr)
library(ggthemr)
library(patchwork)
library(data.table)
library(ggrepel)
ggthemr::ggthemr("pale")
```


# Overview of the problem

The goal is to estimate the relative effect of treatment $A$ vs $B$, noted $\Delta_{t}$. These two treatments have been evaluated in two distinct trials, $AB$ and $AC$. The target estimand is the relative effect of $A$ vs $B$ in the source population of trial $AC$, written $\Delta_{t(AC)}$. 

$\Delta_{t (AB)}$ will be estimated using different estimators: 

- Unadjusted: 
  - Unanchored: naive model
  - Anchored: difference of relative effects $A$ versus $C$ and $B$ versus $C$
- Adjusted
  - Adjusted using a multivariable outcome model (regression model), or univariable weighted outcome model (propensity score)
    - Each of these models will be fitted using either Individual Patient Data (IPD) for both trials, or IPD for the trial $AB$ and AgD for trial $BC$: Population-Adjusted Indirect Comparisons (PAIC). PAIC multivariable outcome model = Simulated Treatment Comparison (STC), univariable weighted outcome model = Matching-Adjusted Indirect Comparisons (MAIC).
    - Each of these models will be either anchored or unanchored.

## Summary of the estimators

|Estimand|Common comparator?|Full IPD?|Estimator|
|--------|------------------|---------|---------|
|Marginal|Yes|Yes|NMA with IPTW PS|
|Marginal|No|Yes|IPTW PS|
|Marginal|Yes|No|Anchored MAIC|
|Marginal|No|No|Unanchored MAIC|
|Conditional|Yes|Yes|IPD NMA|
|Conditional|No|Yes|Regression-based outcome model|
|Conditional|Yes|No|Anchored STC|
|Conditional|No|No|Unanchored STC|



```{r, include=FALSE, warning=FALSE}
source("processing_results.R")
```

\newpage

# Parameters of the different scenarios

- `covariate_names`: Covariates included in the propensity score model
- `outcome_regression_model`: Outcome model used in regression-based estimators
- `model_marginal`: Is the marginal model (propensity score) adequate? Or is it missing variables (confounded), or including non confounding factors variables (imprecise)?
- `model_conditional`: Is the outcome regression model adequate? Or is it missing confoundings, or including non confounding variables?
- `any_nonnormal`: Are at least one of the PF/TEM non-normally distributed?

```{r}
classification_scenario |> 
  select(c(covariate_names, outcome_regression_model, model_marginal, model_conditional, any_nonnormal)) |> 
  mutate(scenario_num = 1:n()) |> 
  select(scenario_num, everything()) |> 
  knitr::kable(caption = "Parameters of different scenarios") |> 
  kableExtra::kable_styling(latex_options = c("scale_down", "hold_position"), 
                            position = "center")
```



Data Generation Model : 

$$Y = \beta_{Y_{X1}} \times X_1 + \beta_{Y_{X2}} \times X_2 + (\beta_{Y_A} + \beta_{Y_{A,X1}} \times X_1 + \beta_{Y_{A,X2}} \times X_2) * A + \beta_{Y_B} \times B + \beta_{Y_C} \times C + \epsilon$$
With $\epsilon \sim N(0, 1)$


<!-- ## Distribution of the estimate  -->

```{r, eval=FALSE}
# Variation of the Estimate
joined_results |>
  ggplot(aes(x = Estimate, y = Adjustment, color = Data, fill = Data)) +
  geom_violin(alpha = 0.8) +  # geom_jitter(alpha = 0.8) +
  facet_grid(Population_parameters_num ~ Estimator_num, scales = "free") +
  theme(legend.position = "bottom")
# +
#   theme_void() +
#   theme(strip.text = element_blank())
  # geom_vline(aes(xintercept = True_effect, linetype = Outcome_type),
  #            show.legend = TRUE) +
  # facet_wrap( ~ Population_parameters_num) +
  # # scale_linetype_manual(values = c("dashed"), labels = c(str_wrap("True conditional effect", width = 10))) +
  # labs(y = NULL,
  #      title = "Treatment effect estimates")
```

<!-- # Standard error of the estimate -->

```{r, eval=FALSE}
joined_results |>
  ggplot(aes(x = Variance, y = Adjustment, color = Data, fill = Data)) +
  geom_violin(alpha = 0.8) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "black") +
  facet_grid(Population_parameters_num ~ Estimator_num, scales = "free") +
  labs(y = NULL,
       x = "Standard error of the estimate",
       title = str_to_title("Treatment effect estimate standard error"))

```


\newpage

# Performance of the estimators

```{r }
# Long indicators
get_bias <- function(obs, theo) {
  mean(obs - theo, na.rm = FALSE)
}
get_RMSE <- function(obs, theo) {
  sqrt(mean((obs - theo)**2, na.rm = FALSE))
}
get_VR <- function(obs, se_obs) {
  mean(se_obs, na.rm = FALSE) / sd(obs, na.rm = FALSE)
}
get_cov_95 <- function(coef, se, theo) {
  ub <- coef + qnorm(0.975)*se
  lb <- coef - qnorm(0.975)*se
  covered <- theo < ub & theo > lb
  mean(covered, na.rm = FALSE)
}

correct_decision <- function(coef, se, theo) {
  ub <- coef + qnorm(0.975)*se
  lb <- coef - qnorm(0.975)*se
  contains_0 <- ub >= 0 & lb <= 0
  sign_estimate_theo_identical <- (theo >= 0 & coef >= 0) | (theo <= 0 & coef <= 0)
  theo_is_null <- theo == 0
  correct_decision <- ifelse(theo_is_null, contains_0, sign_estimate_theo_identical)
  mean(correct_decision, na.rm = FALSE)
}


df_stats <- joined_results |>
  group_by(Population_parameters_num, Estimator_num, Adjustment, Model, Anchored, Data) |>
  summarize(bias = get_bias(Estimate, True_effect ),
            rmse = get_RMSE(Estimate, True_effect ),
            vr = get_VR(Estimate, sqrt(Variance)),
            cov_95 = get_cov_95(Estimate, sqrt(Variance), True_effect ), 
            correct_decision = correct_decision(Estimate, Variance, True_effect),
            .groups = "drop") |>
  pivot_longer(cols = c("bias", "rmse", "vr", "cov_95", "correct_decision"),
               names_to = "indicator", values_to = "values") 
```


```{r, results='asis', fig.height = 7, fig.width = 6, warning = FALSE}
vline_data <- data.frame(
  value = c(0, 0, 1, 0.95), 
  indicator = factor(c("Bias", "RMSE", "VR", "95% coverage"))
)

df_stats |> 
  select(indicator, everything()) |> 
  mutate(Model = ifelse(Model == "Unadjusted", "Unadjusted", toupper(Model)),
         indicator = 
           case_match(indicator, "bias" ~ "Bias", 
                      "rmse" ~ "RMSE", 
                      "vr" ~ "VR", 
                      "cov_95" ~ "95% coverage") |> 
           factor(levels = c("Bias", "RMSE", "VR", "95% coverage"))
  ) |> 
  group_by(Population_parameters_num, Estimator_num) |> 
  group_walk(.f = \(df, df_param) {
    cat(paste0("## Scenario number ", df_param$Estimator_num + (df_param$Population_parameters_num - 1) * 3))
    cat("\n")
    print(knitr::kable(left_join(df_param,
                                 classification_scenario |> 
                                   select(c(population_parameters_num, estimator_num, bY_X1, bY_X2, bY_A_X1, bY_A_X2, f_X1, f_X2, covariate_names, outcome_regression_model)) |>  
                                   rename(regression = outcome_regression_model, 
                                          PS_model = covariate_names),
                                 by = c("Population_parameters_num" = "population_parameters_num", 
                                        "Estimator_num" = "estimator_num")) |> 
                         select(-Population_parameters_num, -Estimator_num), 
                       booktabs = TRUE, 
                       ) |> 
            kableExtra::kable_styling(latex_options = c("scale_down", "hold_position"), 
                                      position = "center")
          )
    print(df |> 
            mutate(combined = factor(paste(Data, Anchored)), 
                   annotation = ifelse(Model == "MAIC", "MAIC", ifelse(Model == "STC", "STC", NA_character_))) |> 
            ggplot(aes(x = values, y = Adjustment, shape = combined, fill = Data, color = Data)) +
            geom_jitter(height = 0.1, size = 3, stroke = 1.5) + 
            facet_wrap("indicator", nrow = 4, scales = "free") +
            scale_fill_manual(values = alpha(c("#de6757", "#eb9050", "#3262ab"), alpha = 0.5)) +
            scale_color_manual(values = alpha(c("#de6757", "#eb9050", "#3262ab"))) +
            scale_shape_manual(values = c(2, 1, 17, 16, 24, 21), guide = "none") +
            geom_vline(aes(xintercept = value), linetype = "dashed", vline_data) +
            geom_text_repel(aes(label = annotation),
                            point.padding = 0.2, 
                            box.padding = 0.3,
                            segment.curvature = -1e-20,
                            arrow = arrow(length = unit(0.015, "npc")), min.segment.length = 0, size = 3) +
            labs(x = NULL, y = NULL) + 
            theme(legend.position = "bottom"))
    cat("\\newpage")
    print(knitr::kable(df |>
                         mutate(values = format(round(values, 3), digits = 3, nsmall = 3, justify = "right")) |>
                         pivot_wider(names_from = indicator, values_from = values)) |> 
            kableExtra::collapse_rows(columns = c(1, 2), latex_hline = "major", row_group_label_position = 'stack'))
    cat("\\newpage")
  })
```

```{r}
knitr::knit_exit()
```


```{r}
ggplot(df_stats, aes(x = values, color = Data, shape = Anchored)) +
  geom_point(aes(y = Adjustment), position = "jitter", size = 2) +
  geom_vline(aes(xintercept = value), linetype = "dashed", data = vline_data) +
  facet_wrap(~ indicator, nrow = 2, scales = "free") +
  theme(legend.position = "bottom")
```


```{r }
bias_plot <- ggplot(filter(df_stats, indicator == "bias"),
                    aes(x = values, color = Data, shape = Anchored)) +
  geom_point(aes(y = Adjustment), position = "jitter", size = 1) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  facet_grid(Population_parameters_num ~ Estimator_num) +
  labs(title = "Bias")
rmse_plot <- ggplot(filter(df_stats, indicator == "rmse"),
                    aes(x = values, color = Data, shape = Anchored)) +
  geom_point(aes(y = Adjustment), position = "jitter", size = 1) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  facet_grid(Population_parameters_num ~ Estimator_num) +
  labs(title = "RMSE")
vr_plot <- ggplot(filter(df_stats, indicator == "vr"),
                    aes(x = values, color = Data, shape = Anchored)) +
  geom_point(aes(y = Adjustment), position = "jitter", size = 1) +
  geom_vline(xintercept = 1, linetype = "dashed") +
  facet_grid(Population_parameters_num ~ Estimator_num) +
  labs(title = "VR")
ci_plot <- ggplot(filter(df_stats, indicator == "cov_95"),
                    aes(x = values, color = Data, shape = Anchored)) +
  geom_point(aes(y = Adjustment), position = "jitter", size = 1) +
  geom_vline(xintercept = 0.95, linetype = "dashed") +
  facet_grid(Population_parameters_num ~ Estimator_num) +
  labs(title = "95% coverage")
```

```{r}
bias_plot
```

```{r}
rmse_plot
```

```{r}
vr_plot
```

```{r}
ci_plot
```

# ```{r }
# bias_plot + rmse_plot + vr_plot + ci_plot +
#   plot_annotation(title = "Performance of the estimators") +
#   plot_layout(guides = "collect") &
#   theme(legend.position = 'bottom',
#         legend.box = "vertical",
#         axis.title = element_blank())
# ```

